R Notebook

### 2022-05-30 final-data-analysis.Rmd
add code

house <- house %>%
  select(- lotSize) %>%     #too many NAs
  mutate(nbhd = replace(nbhd, nbhd == "Battelfield Area", "Battlefield Area"), nbhd = replace(nbhd, nbhd == "princeton Ridge", "Princeton Ridge"), nbhd = replace(nbhd, nbhd == "LIttlebrook", "Littlebrook"), nbhd = replace(nbhd, nbhd == "griggs Farm", "Griggs Farm"), nbhd = replace(nbhd, nbhd == "PrettyBrook Area", "Pretty Brook Area"), nbhd = replace(nbhd, nbhd == "Palmer Sq", "Palmer Square"), nbhd = replace(nbhd, nbhd == "Institute", "Institute Area"), nbhd = replace(nbhd, nbhd == "Rriverside", "Riverside"), nbhd = replace(nbhd, nbhd == "Rriverside", "Riverside"), nbhd = replace(nbhd, nbhd == "Carnegie Lake", "Carnegie Lake Area"), nbhd = replace(nbhd, nbhd == "Town", "Township")) %>%     #fix area names
  filter(! nbhd %in% c("Edgerstoune", "Northridge", "Rushbrook", "Russell Estates", "The Preserve", "Governors Lane", "Princeton Borough", "Preserve", "Queenston")) %>%  #delete areas with few houses
  filter(! style %in% badStyles$style) %>%  #delete styles with at most 3 houses
  mutate(style = replace(style, style == "Bi-level", "Bi-Level"), style = replace(style, style == "End Unit", "End-Unit"), style = replace(style, style == "Townhouse", "Townhome"))  #fix style names

@@

houseNorm <- house %>%
  mutate_at(c("bed", "fullBath", "halfBath", "age", "originalPrice", "lastPrice", "monthSold", "daySold", "yearSold", "marketDays"), ~(scale(.) %>% as.vector))

@@

eval_metrics <- function(model, df, predictions, target){
    resids = df[,target] - predictions
    resids2 = resids**2
    N = length(predictions)
    r2 = round(summary(model)$r.squared, 3)
    adj_r2 = round(summary(model)$adj.r.squared, 3)
    rmse = round(sqrt(sum(resids2)/N), 3)
    return(c(adj_r2, rmse))
}

@@

linearModel <- function(df, testSet) {
  lr <- lm(soldPrice ~ ., data = filter(df, id != testSet))
  summary(lr)
  predictions <- predict(lr, newdata = filter(df, id == testSet))
  eval_metrics(lr, filter(df, id == testSet), predictions, target = "soldPrice")
}

@@

subsetModel <- function(df, testSet) {

}

performance(houseSplit, linearModel)
```

>>>>

house <- house %>%
  select(- lotSize) %>%     #too many NAs
  select(- originalPrice) %>%
  select(- lastPrice) %>%  #we omit these two predictors because they are too similar to the response variable 
  mutate(nbhd = replace(nbhd, nbhd == "Battelfield Area", "Battlefield Area"), nbhd = replace(nbhd, nbhd == "princeton Ridge", "Princeton Ridge"), nbhd = replace(nbhd, nbhd == "LIttlebrook", "Littlebrook"), nbhd = replace(nbhd, nbhd == "griggs Farm", "Griggs Farm"), nbhd = replace(nbhd, nbhd == "PrettyBrook Area", "Pretty Brook Area"), nbhd = replace(nbhd, nbhd == "Palmer Sq", "Palmer Square"), nbhd = replace(nbhd, nbhd == "Institute", "Institute Area"), nbhd = replace(nbhd, nbhd == "Rriverside", "Riverside"), nbhd = replace(nbhd, nbhd == "Rriverside", "Riverside"), nbhd = replace(nbhd, nbhd == "Carnegie Lake", "Carnegie Lake Area"), nbhd = replace(nbhd, nbhd == "Town", "Township")) %>%     #fix area names
  filter(! nbhd %in% c("Edgerstoune", "Northridge", "Rushbrook", "Russell Estates", "The Preserve", "Governors Lane", "Princeton Borough", "Preserve", "Queenston")) %>%  #delete areas with few houses
  filter(! style %in% badStyles$style) %>%  #delete styles with at most 3 houses
  mutate(style = replace(style, style == "Bi-level", "Bi-Level"), style = replace(style, style == "End Unit", "End-Unit"), style = replace(style, style == "Townhouse", "Townhome"))  #fix style names

@@

houseNorm <- house %>%
  mutate_at(c("bed", "fullBath", "halfBath", "age", "monthSold", "daySold", "yearSold", "marketDays"), ~(scale(.) %>% as.vector))

@@

evalMetrics <- function(model, df, predictions, target){
    resids = df[,target] - predictions
    resids2 = resids**2
    N = length(predictions)
    r2 = round(summary(model)$r.squared, 3)
    adjR2 = round(summary(model)$adj.r.squared, 3)
    rmse = round(sqrt(sum(resids2)/N), 3)
    return(c(adjR2, rmse))
}

@@

linearModel <- function(df, testSet) {
  lr <- lm(soldPrice ~ ., data = select(filter(df, id != testSet), -id))
  summary(lr)
  predictions <- predict(lr, newdata = select(filter(df, id == testSet), -id))
  evalMetrics(lr, select(filter(df, id == testSet), -id), predictions, target = "soldPrice")
}

@@

forwardModel <- function(df, testSet) {
  fitNone <- lm(soldPrice ~ 1 , data = select(filter(df, id != testSet), -id))
  forward <- step(fitNone, direction = "forward", trace = 0, scope = formula(lm(soldPrice ~ ., data = select(filter(df, id != testSet), -id))))
  predForward <- predict(forward, newdata = select(filter(df, id == testSet), -id))
  evalMetrics(forward, select(filter(df, id == testSet), -id), predForward, target = "soldPrice")
}

backwardModel <- function(df, testSet) {
  fitAll <- lm(soldPrice ~ ., data = select(filter(df, id != testSet), -id))
  backward <- step(fitAll, direction = "backward", trace = 0)
  predBackward <- predict(backward, newdata = select(filter(df, id == testSet), -id))
  evalMetrics(backward, select(filter(df, id == testSet), -id), predBackward, target = "soldPrice")
}

performance(houseSplit, linearModel)
performance(houseSplit, forwardModel)
performance(houseSplit, backwardModel)
```

```{r}
#we will use the glmnet package to build regularized regression models
#these include ridge, lasso, and elastic net

# install.packages("glmnet")
# install.packages("caret")
library(glmnet)
library(caret)

dummies <- dummyVars(soldPrice ~ ., data = houseSplit[,colnames(houseNorm)])
```
